Pulsar Acceleration Shifts from nearby Supernova Explosion

### 2016-04-29 draft3_1C.tex

In Sec.\ref{sec-3d}, we study the general case in which the pulsar and SN are not aligned. The above acceleration change happens when the shell hits the pulsar. However, before seeing that, we will first see the SN explosion, and thereafter the pulses must cross the neutrino shell before reaching us. Such shell-crossing also induces observable changes in the pulsar period at similar order of magnitudes \cite{OluPie13}. Combining these two effects, we derive the a unique signature in pulsar observation accompanying the SN explosion.

@@

Before $t_1$, both the pulsar and the earth are in metric $i$. After $t_2$, both are in metric $f$. Thus we simply have
\begin{eqnarray}
\frac{\dot{w}_e}{w_e} &=& -\frac{M_i}{r_p^2}\cos\theta~, \ \ \ {\rm for} \ \ t<t_1~,
\label{eq-before} \\ \nonumber
&=& -\frac{M_f}{r_p^2}\cos\theta~, \ \ \ {\rm for} \ \ t>t_2 = t_1 + r_p(1-\cos\theta)~,
\end{eqnarray}
which comes from the projection of gravitational acceleration in our direction. Between $t_1$ and $t_2$, a pulse originates in metric $i$, but must cross the neutrino shell (being caught-up by it) and enter metric $f$ before reaching the earth. This shell-crossing leads to a continuous change in the observed frequency. The change in photon frequency when the pulsar remains stationary was calculated in \cite{OluPie13}. Here we include the fact that before $t_2$, the pulsar is still feeling the same acceleration which lead to Eq.~(\ref{eq-before}). Without loss of generality, we set $t_1=0$, then we can combine with the result of \cite{OluPie13} to get the full change in frequency through out the process.
\begin{eqnarray}
w_e(t) &=& w_0 \left[ 1 -  t\frac{M_i}{r_p^2}\cos\theta \right]~, 
\ \ \ {\rm for} \ \ t < 0~, \label{eq-beforeW}  \\ \nonumber
&=& w_0
\left[ 1 -  t\frac{M_i}{r_p^2}\cos\theta + \frac{4\delta M t^3}{(t^2 + r_p^2\sin^2\theta)^2} \right]~, 
\ \ \ {\rm for} \ \ 0 < t <  r_p(1 - \cos\theta)~, \\ \nonumber
&=& w_0
\left[ 1-\frac{M_i}{r_p}\cos\theta(1-\cos\theta) + \frac{\delta M}{r_p}(1-\cos\theta) -  \bigg(t-r_p(1-\cos\theta)\bigg) \frac{M_f}{r_p^2}\cos\theta \right]  \\ \nonumber
& & {\rm for} \ \ t > r_p(1-\cos\theta)~.
\end{eqnarray}
Here $w_0 = w_e(0)$ is the frequency observed on earth at the moment when we see the SN explosion.

@@

Beyond this simple estimation, the detail shape of the time-dependent change of the arrival time is given by integrating Eq.~(\ref{eq-beforeDW}). As already shown in \cite{OluPie13}, the first stage of such change depends only on the ``impact parameter'' $r_p\sin\theta$. This corresponds to the angular distance between the pulsar and SN in our sky, which can be measured on its own quite accurately. On the other hand, the discontinuous change in the derivative of frequency happens at $t=r_p(1-\cos\theta)$. That corresponds to the depth difference between SN and pulsar, which is notoriously hard to measure. We select a few values of $\theta$ and plot their actual $\delta T_{\rm arrival}$ as a function of time in Fig.\ref{fig-3x3}. With a good measure, it should not be difficult to fit these curves, therefore an improved measurement of their depth difference.

>>>>

In Sec.\ref{sec-3d}, we study the general case in which the pulsar and SN are not aligned. The above acceleration change happens when the shell hits the pulsar. However, before seeing that, we will first see the SN explosion, and thereafter the pulses must cross the neutrino shell before reaching us. Such shell-crossing also induces observable changes in the pulsar period at similar order of magnitudes \cite{OluPie13}. Combining these two effects, we derive a unique signature in pulsar observation accompanying the SN explosion.

@@

Before $t_1$, both the pulsar and the earth are in metric $i$. After $t_2$, both are in metric $f$. Thus we simply have
\begin{eqnarray}
\frac{\dot{w}_e}{w_e} &=& -\frac{M_i}{r_p^2}\cos\theta~, \ \ \ {\rm for} \ \ t<t_1~,
\label{eq-before} \\ \nonumber
&=& -\frac{M_f}{r_p^2}\cos\theta~, \ \ \ {\rm for} \ \ t>t_2 = t_1 + r_p(1-\cos\theta)~,
\end{eqnarray}
which comes from the projection of gravitational acceleration in our direction. Between $t_1$ and $t_2$, a pulse originates in metric $i$, but must cross the neutrino shell (being caught-up by it) and enter metric $f$ before reaching the earth. This shell-crossing leads to a continuous change in the observed frequency. The change in photon frequency when the pulsar remains stationary was calculated in \cite{OluPie13}. Here we include the fact that before $t_2$, the pulsar is still feeling the same acceleration which lead to Eq.~(\ref{eq-before}). Without loss of generality, we set $t_1=0$, then we can combine with the result of \cite{OluPie13} to get the full change in frequency through out the process.
\begin{eqnarray}
w_e(t) &=& w_0 \left[ 1 -  t\frac{M_i}{r_p^2}\cos\theta \right]~, 
\ \ \ {\rm for} \ \ t < 0~, \label{eq-beforeW}  \\ \nonumber
&=& w_0
\left[ 1 -  t\frac{M_i}{r_p^2}\cos\theta + \frac{4\delta M t^3}{(t^2 + r_p^2\sin^2\theta)^2} \right]~, 
\ \ \ {\rm for} \ \ 0 < t <  r_p(1 - \cos\theta)~, \\ \nonumber
&=& w_0
\left[ 1-\frac{M_i}{r_p}\cos\theta(1-\cos\theta) + \frac{\delta M}{r_p}(1-\cos\theta) -  \bigg(t-r_p(1-\cos\theta)\bigg) \frac{M_f}{r_p^2}\cos\theta \right]  \\ \nonumber
& & {\rm for} \ \ t > r_p(1-\cos\theta)~.
\end{eqnarray}

Here $w_0 = w_e(0)$ is the frequency observed on earth at the moment when we see the SN explosion. 
In the second line, the final term is what is calculated\footnote{See Eq (23) in \cite{OluPie13}. We have replaced $b$ in that equation with $r_p \sin \theta$ to be consistent with notation used in this paper.} in \cite{OluPie13} and this incorporates the effect of the pulses crossing the neutrino shell. In going to the final line we have simply set $t = r_p (1 - \cos \theta)$ as the effects in the second line will build up to that time and then will remain constant. The final term in the third line is just the acceleration of the pulsar in the new metric after shell crossing.

@@

Beyond this simple estimation, the detail shape of the time-dependent change of the arrival time is given by integrating Eq.~(\ref{eq-beforeDW}). As already shown in \cite{OluPie13}, the first stage of such change depends only on the ``impact parameter'' $r_p\sin\theta$. This corresponds to the angular distance between the pulsar and SN in our sky, which can be measured on its own quite accurately. On the other hand, the discontinuous change in the derivative of frequency happens at $t=r_p(1-\cos\theta)$ as the pulsar is now in the final metric and so feels a different acceleration. Note that $t = r_p(1-\cos \theta)$ is also the time that corresponds to the depth difference between SN and pulsar which is notoriously hard to measure. We select a few values of $\theta$ and plot their $\delta T_{\rm arrival}$ as a function of time in Fig.\ref{fig-3x3}. With a good measure, it should not be difficult to fit these curves, therefore an improved measurement of their depth difference.
